### 2015-10-17 doc/paper.tex
Cleanup

\documentclass[11pt]{amsart}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath, stmaryrd, scalerel}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{proof}
\usepackage{supertabular}
\usepackage{unicode-math}
\usepackage{xltxtra}
\usepackage{xunicode}

\usepackage{setspace,url,hyperref}

% fonts
\setromanfont[Mapping=tex-text, Ligatures=Rare, Numbers=OldStyle]{Linux Libertine O}
\setmathfont{Asana Math}

\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[thm]{Definition}
\newtheorem{example}[thm]{Example}
\theoremstyle{remark}
\newtheorem{remark}[thm]{Remark}
\numberwithin{equation}{section}

\lstset{
  language=[Objective]Caml,
  extendedchars=\true,
  inputencoding=utf8,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  basicstyle=\ttfamily,
  xleftmargin=6mm,
  framesep=4mm,
  moredelim=*[s][\itshape]{(*}{*)},
  moredelim=[is][\textcolor{darkgray}]{§}{§},
  escapechar=°,
  keywords=[6]{signature, structure, struct type, sig, end},
  keywordstyle=\color{MidnightBlue},
  stringstyle=\color[rgb]{0.545098, 0.278431, 0.364706},
  commentstyle=\color[rgb]{0.698039, 0.133333, 0.133333},
  numberstyle=\color[rgb]{0.372549, 0.619608, 0.627451},
  literate=
    {->}{{$\rightharpoonup$}}{1}
    {*}{{$\otimes$}}{1}
}

\def\InputModeColorName{MidnightBlue}
\def\OutputModeColorName{Maroon}
\newcommand\InputMode[1]{{\color{\InputModeColorName}{#1}}}
\newcommand\OutputMode[1]{{\color{\OutputModeColorName}{#1}}}
\newcommand\Quote[1]{{\color{Gray}{«}}#1{\color{Gray}{»}}}

\newcommand\HypJ[2]{#1\ \ (#2)}
\newcommand\GenJ[2]{\vert_{\InputMode{#1}}\; #2}
\newcommand\JJ{J}
\newcommand\Tuple[1]{\left\langle#1\right\rangle}
\newcommand\HTuple[1]{\left\lBrace#1\right\rBrace}
\newcommand\RSet{\mathfrak{R}}
\newcommand\JSet{\mathfrak{J}}

\newcommand\IsTac[1]{\InputMode{#1}\ \textit{tactic}}

%% subgoals
\newcommand\SG[2]{\InputMode{#1}\,{\color{Black}:}\,\OutputMode{#2}}

%% refinement
\newcommand\Refine[4]{\InputMode{#1}\Vdash\InputMode{#2}\Rightarrow\OutputMode{#4}\dashv\OutputMode{#3}}
\newcommand\NoRefine[2]{\InputMode{#1}\Vdash\InputMode{#2}\Uparrow}

\newcommand\Member[2]{\InputMode{#1}\in\InputMode{#2}}
\newcommand\IdTac{\mathtt{id}}
\newcommand\FailTac{\mathtt{fail}}
\newcommand\OrElseTac[2]{#1\vert#2}
\newcommand\ThenTac[2]{#1;#2}
\newcommand\ThenLTac[2]{#1;[#2]}
\newcommand\ProdIntroRule{\mathtt{ProdIntro}}
\newcommand\DProdIntroRule[1]{\mathtt{DepProdIntro}\{#1\}}
\newcommand\DProdIntroRuleNullary{\mathtt{DepProdIntro}}
\newcommand\TyProd[2]{#1 \times #2}
\newcommand\TyDProd[3]{(#2:#1) \times #3}
\newcommand\TyMember[2]{#1\in#2}
\newcommand\Dom[1]{\left\vert#1\right\vert}
\newcommand\SubsetEq[2]{\InputMode{#1}\subseteq\InputMode{#2}}
\newcommand\FV[1]{\textit{FV}\left(#1\right)}
\newcommand\Subst[3]{\left[#1\;/\;#2\right]#3}
\newcommand\ModLCF{\textbf{\emph{Modernized LCF}}}
\newcommand\DepLCF{\textbf{\emph{Dependent LCF}}}

\begin{document}

\title{Dependency in Refinement Logics}

\author{Jonathan Sterling}
\address{}
\maketitle

\begin{abstract}

  We propose a new generalization of LCF-style refinement logics called \DepLCF,
  which allows the definition of refinement rules which express a dependency
  between subgoals, without the use of unification variables.

\end{abstract}

\onehalfspacing

Over the past four decades, there have been many incarnations of the \emph{LCF}
(Logic for Computable Functions) interface, but the one that will concern us
here is its extension to \emph{LCF with Validations} as found in Cambridge LCF,
which outfits the proving activity with the synthesis of explicit evidence.
The Cambridge LCF signature is as follows:

%% cite "Logic and Computation: Interactive Proof with Cambridge LCF"

\begin{lstlisting}
  signature CAMBRIDGE_LCF =
  sig
    type form
    type thm
    type proof = thm list -> thm
    type goal = form list * form
    type tactic = goal -> goal list * proof
  end
\end{lstlisting}

The \verb!goal! type represents sequents, where \verb!form! represents logical
propositions. The LCF methodology, however, may be used to give a refinement
treatment to many kinds of judgments, not just sequents; therefore, it
is better to not include this in the core signature at all, and make the
type of judgments abstract; for clarity, we will replace the name \verb!thm!
with \verb!evidence!.

\begin{lstlisting}
  signature LCF =
  sig
    type judgment
    type evidence
    type proof = evidence list -> evidence
    type tactic = judgment -> judgment list * proof
  end
\end{lstlisting}

Then, a tactic is something that refines a judgment to a list of judgments (its
subgoals), and synthesizes its evidence (provided the evidence of its
subgoals). There are many different tactics which can be implemented
generically over this signature; here are a few:

\begin{lstlisting}
  signature TACTICALS =
  sig
    structure Lcf : LCF
    val ID : Lcf.tactic
    val FAIL : Lcf.tactic
    val THEN : Lcf.tactic * Lcf.tactic -> Lcf.tactic
    val THENL : Lcf.tactic * Lcf.tactic list -> Lcf.tactic
  end
\end{lstlisting}

\section{modernized lcf: the logic of tactics}

In order to study the design space for LCF refiners, we would like to give a
judgmental characterization of tactic systems, which we will call $\ModLCF$.
To begin with, note that the type of validations (\verb!proof!) in the ML
implementation is is essentially a HOAS (higher-order abstract syntax) encoding
of a hypothetical proof or synthesis of a judgment.  With this insight in hand,
we are in a position to unify the list of subgoals and the validation generated
by a tactic into a single concept, namely that of a \emph{hypothetical proof}
$E$ whose free variables are explained in a context $\Psi$ of subgoals.

To make the preceding observations precise, we can characterize the behavior of
a \ModLCF\ refiner judgmentally via two forms of judgment,
\framebox{$\Refine{\JJ}{\tau}{\Psi}{E}$} and \framebox{$\NoRefine{\JJ}{\tau}$}, where $\JJ$ is a judgment of the
logical theory, $\tau$ is a tactic, $\Psi$ is a context of judgments,
representing the subgoals generated by the tactic $\tau$, and $E$ is the
synthesis of the judgment $\JJ$, binding variables $\Dom{\Psi}$ which represent
the syntheses of the subgoals.

The meaning of $\Refine{\JJ}{\tau}{\Psi}{E}$ is that $\tau$ is applicable to
demonstrating the judgment $\JJ$, producing synthesis $e$ under the assumptions
that the judgments in $\Psi$ can be demonstrated. The divergence judgment
$\NoRefine{\JJ}{\tau}$ expresses the inapplicability of $\tau$ to $\JJ$.  In
practice, we will explain only the assertion conditions for one of
$\Refine{\JJ}{\tau}{\Psi}{E}$ and $\NoRefine{\JJ}{\tau}$, and implicitly take
the other to be its complement.

\begin{remark}
  The refinement judgments $\Refine{\JJ}{\tau}{\Psi}{E}$ and
  $\NoRefine{\JJ}{\tau}$ are not higher-order judgments, because the variable
  $\JJ$ ranges not over judgments of the refinement theory, but of the
  object theory.
\end{remark}

Let $\JSet$ be the open-ended collection of judgments in our logical theory,
and let $\RSet$ be a collection of rule names. Each rule $\Member{R}{\RSet}$
must be interpretable as a tactic, i.e.\ the meaning of the assertions
$\Refine{\JJ}{R}{\Psi}{E}$ and $\NoRefine{\JJ}{R}$ must be explained for
$\Member{J}{\JSet}$. We say $\IsTac{\tau}$ in case for all object-judgments
$\Member{\JJ}{\JSet}$, the assertion conditions for
$\Refine{\JJ}{\tau}{\Psi}{E}$ and $\NoRefine{\JJ}{\tau}$ are disjoint, and
moreover, if $\Refine{\JJ}{\tau}{\Psi}{E}$, then
$\SubsetEq{\FV{E}}{\Dom{\Psi}}$.

Numerous general purpose tactics can be defined over a logical theory,
including identity, failure, disjunction and sequencing:

\begin{gather*}
  \infer{
    \Refine{\JJ}{\IdTac}{\SG{\alpha}{\JJ}}{\alpha}
  }{
  }
  \qquad
  \infer{
    \NoRefine{\JJ}{\FailTac}
  }{
  }\\[6pt]
  \infer{
    \Refine{\JJ}{\OrElseTac{\tau_1}{\tau_2}}{\Psi}{E}
  }{
    \Refine{\JJ}{\tau_1}{\Psi}{E}
  }
  \qquad
  \infer{
    \Refine{\JJ}{\OrElseTac{\tau_1}{\tau_2}}{\Psi}{E}
  }{
    \NoRefine{\JJ}{\tau_1} &
    \Refine{\JJ}{\tau_2}{\Psi}{E}
  }\\[6pt]
  \infer{
    \Refine{\JJ}{\ThenTac{\tau_1}{\tau_2}}{\bigoplus_{\Dom\Phi}\Psi}{\Subst{F}{\Dom\Phi}{E}}
  }{
    \Refine{\JJ}{\tau_1}{\Phi}{E} &
    \GenJ{\alpha}{
      \HypJ{
        \Refine{\Phi(\alpha)}{\tau_2}{\Psi_\alpha}{F_\alpha}
      }{
        \Member{\alpha}{\Dom\Phi}
      }
    }
  }
\end{gather*}

\begin{remark}
  The nominal treatment that we have given here allows for a much more
  economical presentation of the standard tacticals, which are quite ardruous
  to define in the HOAS treatment used in ML implementations.
\end{remark}

\begin{thm}
  The above rules all define valid tactics:.
  \begin{enumerate}
    \item $\IsTac{\IdTac}$.
    \item $\IsTac{\FailTac}$.
    \item If $\IsTac{\tau_1}$ and $\IsTac{\tau_2}$, then
      $\IsTac{\OrElseTac{\tau_1}{\tau_2}}$.
    \item If $\IsTac{\tau_1}$ and $\IsTac{\tau_2}$, then
      $\IsTac{\ThenTac{\tau_1}{\tau_2}}$.
  \end{enumerate}
\end{thm}
\begin{proof}
  It suffices to verify that the synthesis $E$ of each tactical is well-scoped
  in $\Psi$; in each case, this follows by induction.
  \begin{enumerate}
    \item[(1--2)] Immediate.
    \item[(3)]
      By induction on derivations of
      $\Refine{\JJ}{\OrElseTac{\tau_1}{\tau_1}}{E}{\Psi}$.
      \begin{description}
        \item[Case $\Refine{\JJ}{\tau_1}{\Psi}{E}$] Validity follows from the
          inductive hypothesis $\IsTac{\tau_1}$.
        \item[Case $\NoRefine{\JJ}{\tau_1}$] Validity follows from the
          inductive hypothesis $\IsTac{\tau_2}$.
      \end{description}

    \item[(4)] We need to show that
      $\SubsetEq{\FV{\Subst{F}{\Dom\Phi}{E}}}{\Dom{\bigoplus_{\Dom\Phi}\Psi}}$.
      From our inductive hypotheses, it is evident that
      $\SubsetEq{\FV{E}}{\Dom\Phi}$ and that for each $\Member{\alpha}{\Phi}$,
      $\SubsetEq{\FV{F_\alpha}}{\Dom{\Psi_\alpha}}$; all of the free variables
      of the term got by substituting each $F_\alpha$ for
      $\Member{\alpha}{\Dom\Phi}$ clearly reside in one of the fibres of the
      family of contexts $\Psi$, and so they must comprise a subset of the
      union $\bigoplus_{\Dom\Phi}\Psi$ of $\Psi$'s fibres.

  \end{enumerate}
\end{proof}

\end{document}